Spitzer Sgr target selection

### 2013-11-08 text/notes/spitzer_target_selection/spitzer_sgr.tex
target selection

@@

Next we take 30 bins in $\Lambda$ and compute the median and $\pm$3-$\sigma$ distances for each bin from the LM10 particle data for the leading and trailing arm separately. We then find any star that could feasibly be within $\pm$3-$\sigma$ of the median distance to each arm, taking into account the distance errors. Figure~\ref{fig:lm10_selection} shows these (soft) margins overlaid with the LM10 particles, while figure~\ref{fig:css_selection} shows the CSS RRLs with RVs selected with this method. We ignore trailing arm data with $\Lambda>180$ since Belokurov et al. 2013 show with BHB stars that the trailing arm may not match the LM10 model past apocenter.

We then do the same thing with $v_{\rm gsr}$ by taking 30 bins in $\Lambda$ and repeating the above algorithm. Figure~\ref{fig:lm10_vgsr} shows the median and $\pm$3-$\sigma$ velocities for these bins, and figure~\ref{fig:css_vgsr} shows the CSS RRLs selected with this window. The remaining sample consists of 193 RR Lyrae, with a total required integration time of $\sim$1000 hr --- way over the budgeted time. We 1) want to get precise distances along many phases of the stream and 2) don't want to spend all of our time on distant targets which require the most exposure time. To select a subset, we take 5 stars at random from 15 degree bins in $\Lambda$. The final sample is shown in red in Figure~\ref{fig:final_sample}, with the other likely stream members in black, overplotted on the LM10 particles.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{lm10_selection.pdf}
\caption{ LM10 particles from each Sgr arm, with median distance (red) and $\pm$3-$\sigma$ (green) distance values for 30 bins in $\Lambda$. }\label{fig:lm10_selection}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{css_selection_vgsr.pdf}
\caption{ CSS RRLs with measured radial velocities selected to be within $\pm$3-$\sigma$ + distance error from the median LM10 distance in each bin from fig.~\ref{fig:lm10_selection}. }\label{fig:css_selection}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{lm10_vgsr_selection.pdf}
\caption{ LM10 particles from each Sgr arm, with median radial velocity (red) and $\pm$3-$\sigma$ (green) values for 30 bins in $\Lambda$. }\label{fig:lm10_vgsr}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{css_selection_from_vgsr.pdf}
\caption{ CSS RRLs selected to be within $\pm$3-$\sigma$ from the median LM10 radial velocity in each bin from fig.~\ref{fig:lm10_vgsr}. }\label{fig:css_vgsr}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{final_sample.pdf}
\caption{ Final sample of CSS RRLs selected to be observed with Spitzer (red) and other RRLs that meet the criteria, but are unable to observe due to time constraints. Concentric circles with labels show required total integration time (over 12 visits) for a typical RR Lyrae at that distance. }\label{fig:final_sample}
\end{center}
\end{figure}

>>>>

\section{An overview of the sample}
\begin{itemize}
	\item Three clumps (29 stars total) along the leading tail (sampling in orbital phase);
	\item a clump of 8 stars selected to be associated with the bifurcated part of the stream;
	\item a clump of 10 stars selected to be associated with the nearby part of the trailing arm;
	\item 16 stars selected from the trailing arm with $\Lambda$ $<$ 180.
\end{itemize}
Though the bifurcated stream targets and nearby trailing targets are selected based on distance, radial velocity, and $B$, it is possible that they will not reveal anything (still interesting) -- these two components only account for $\sim$15\% of the time, so we think this is acceptable.

\section{Details about target selection}

@@

Next we take 30 bins in $\Lambda$ and compute the median and $\pm$3-$\sigma$ distances for each bin from the LM10 particle data for the leading and trailing arm separately. We then find any star that could feasibly be within $\pm$3-$\sigma$ of the median distance to each arm, taking into account the distance errors. Figure~\ref{fig:dist_selection} shows these margins overlaid with the CSS RRLs selected from both the leading and trailing arm to match in distance. We manually offset the leading arm by $-$5~kpc where necessary to account for the apparent shift in distance in the CSS overdensity. We then do the same thing with $v_{\rm gsr}$ by taking 30 bins in $\Lambda$ and repeating the above algorithm. Figure~\ref{fig:vgsr_selection} shows the CSS RRLs selected with this window.

Figure~\ref{fig:koposov} (Figure~1 from Koposov et al. (2012)) shows the bifurcated part of the stream and main overdensity observed in SDSS and how they each vary in $\Lambda$ vs. $B$. For $180 < \Lambda < 300$, the main leading arm is actually below $B=0$ and the bifurcation is at or slightly above the midplane. The opposite is true for $\Lambda < 180$. Given this, we only select leading arm targets with $B<0$ for large $\Lambda$, and $B<5$ elsewhere. Figure~\ref{fig:LB} shows in green the targets selected to be associated with the bifurcation, and in blue all others from the far wraps.

The remaining sample consists of 66 RR Lyrae, with a total required integration time of $\sim$280 hr and is shown in Figure~\ref{fig:final_sample} (green), along with other RR Lyrae that meet the above cuts but are not in the chosen clumps (red), overlaid with the LM10 particles.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{dist_selection.pdf}
\caption{  CSS RRLs selected to be within $\pm$3-$\sigma$ from the median LM10 distance in each bin. }\label{fig:dist_selection}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{vgsr_selection.pdf}
\caption{ CSS RRLs selected to be within $\pm$3-$\sigma$ from the median LM10 radial velocity in each bin. }\label{fig:vgsr_selection}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{koposov.png}
\caption{ Figure~1 from Koposov et al. (2012), showing Sgr in the SDSS footprint. The bifurcation is near $B=0$ for $\Lambda>180$ and near $B=10$ for $\Lambda<180$. }\label{fig:koposov}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{L_vs_B.pdf}
\caption{ $\Lambda$ vs. $B$ for LM10 particles (black) and selected particles (excluding the nearby trailing stars, for now). Blue are selected to not be associated with the bifurcation, green are the clump selected to be associated with the bifurcation. }\label{fig:LB}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{with_near_trailing_xz.pdf}
\caption{ Final sample of CSS RRLs selected to be observed with Spitzer (green) and other RRLs that meet the criteria, but are unable to observe due to time constraints (red). }\label{fig:final_sample}
%Concentric circles with labels show required total integration time (over 12 visits) for a typical RR Lyrae at that distance.
\end{center}
\end{figure}
